---
title: "MA615 HW#2"
author: "Xinqiao Zhang"
date: "10/2/2019"
output: pdf_document
---
```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, include=FALSE)
pacman::p_load("tidyverse","knitr","kableExtra")
```
## Problem 1
Load the gapminder data from the gapminder package.
```{r}
library("gapminder")
```
How many continents are included in the data set?
```{r}
unique(gapminder$continent)
```
\qquad There are 5 continents included in the dataset.   
How many countrys are included? How many countries per continent?
```{r}
unique(gapminder$country)
```
\qquad There are 142 countries in the dataset, so it is 28.4 countries per continent on average.  
Using the gapminder data, produce a report showing the continents in the dataset, total population per continent, and GDP per capita. Be sure that the table is properly labeled and suitable for inclusion in a printed report.
```{r, include=TRUE}
gap_1 <- gapminder %>% group_by(year=2007, continent) %>% summarize(ttl_pop= sum(as.numeric(pop)), GDPperCap=sum(gdpPercap))
gap_1
```
\qquad Since the data contains data from different years, it would be reasonbale to summarize the entries by years. Since there are too many years to show, we only show the data from the latest year, 2007.    
Produce a well-labeled table that summarizes GDP per capita for the countries in each continent, contrasting the years 1952 and 2007.
```{r, include=TRUE}
gap_1952<- gapminder %>% filter(year==1952) %>% group_by(continent) %>% summarize(GDPperCap=sum(gdpPercap))
gap_2007<- gapminder %>% filter(year==2007) %>% group_by(continent) %>% summarize(GDPperCap=sum(gdpPercap))

kable(gap_1952, caption="GDP per Capital in 1952") %>% 
  kable_styling(latex_option="striped")
kable(gap_2007, caption="GDP per Capital in 2007") %>% 
  kable_styling(latex_option="striped")
```
Product a plot that summarizes the same data as the table. There should be two plots per continent.
*GDP in 1952*
```{r, include=TRUE}
gap_1952t<- gapminder %>% filter(year==1952) %>% group_by(continent, country) %>% summarize(gdpPercap)
ggplot(gap_1952t)+ aes(x=country, weights=gdpPercap) + geom_bar() + xlab("country") + ylab("gdpPercap") 
```
*GDP in 2007*
```{r, include=TRUE}
gap_2007t<- gapminder %>% filter(year==2007) %>% group_by(continent, country) %>% summarize(gdpPercap)
ggplot(gap_2007t)+ aes(x=country, weights=gdpPercap) + geom_bar() + xlab("country") + ylab("gdpPercap") 
```
Which countries in the dataset have had periods of negative population growth?
Illustrate your answer with a table or plot.
```{r, include=TRUE}
gap_grow<- gapminder %>% filter(year==1952 | year==2007) %>% group_by(continent, country) %>% summarize(growth=diff(gdpPercap))
gap_neg<- gap_grow %>% filter(growth<0)
kable(gap_neg, caption = "Countries with negative growth in GDP")
```
\qquad The countries listed above had negative growth in the period.   
Which countries in the dataset have had the highest rate of growth in per capita GDP?
Illustrate your answer with a table or plot.
```{r, include=TRUE}
max_growth<- gap_grow %>% arrange(growth)
kable(max_growth[138:142,], caption="5 Countries with the highest growth in GDP")
```
\qquad According to the table, Singapore, Asia has the highest growth in GDP from 1952 to 2007.   
## Problem 2
The data for Problem 2 is the Fertility data in the AER package. This data is from the 1980 US Census and is comprised of date on married women aged 21-35 with two or more children. The data report the gender of each woman’s first and second child, the woman’s race, age, number of weeks worked in 1979, and whether the woman had more than two children.  
```{r}
library("AER")
data(Fertility)
```
There are four possible gender combinations for the first two Children. Product a plot the contracts the frequency of these four combinations. 
```{r, include=TRUE}
child_comb<- Fertility %>% group_by(gender1,gender2) %>% summarize(moms=length(age))
child_comb<- child_comb %>% unite(combination, gender1, gender2)
kable(child_comb)
ggplot(child_comb) + 
  geom_bar(mapping = aes(x = combination, y = moms), stat = "identity")
```
Are the frequencies different for women in their 20s and wemen who are older than 29?
```{r}
moms_30minus <- Fertility %>% filter(age<=29) %>% group_by(gender1,gender2) %>% summarise(moms=length(age))
moms_30minus <- moms_30minus %>% unite(combination, gender1, gender2)
kable(moms_30minus)
ggplot(moms_30minus) + 
  geom_bar(mapping = aes(x = combination, y = moms), stat = "identity")

moms_30plus <- Fertility %>% filter(age>29) %>% group_by(gender1,gender2) %>% summarise(moms=length(age))
moms_30plus <- moms_30plus %>% unite(combination, gender1, gender2)
kable(moms_30plus)
ggplot(moms_30plus) + 
  geom_bar(mapping = aes(x = combination, y = moms), stat = "identity")
```
\qquad From the bar chart we can tell that the frequencies of the four combinations of moms under 29 are obviously less the the frequencies of those above 29. However, the rough relationship of the four combinations does not vary too much. The frequencies of having two males children are leading in both groupds.   
Produce a plot that contrasts the frequency of having more than two children by race and ethnicity.
```{r, include=TRUE}
moms_race<- Fertility %>% group_by(afam,hispanic,other) %>% summarize(moms=sum(morekids=="yes"))
moms_race$"Ethnicity" = c("Caucasian","Other","Hispanic","Hispanic and Other","African-American","African-American and Hispanic")
ggplot(moms_race) +
 aes(x = Ethnicity, weight = moms) +
 geom_bar() +
  ylab("Frequency of having more than two kids")
```

## Problem 3
Use the mtcars and mpg datasets.
```{r}
data(mtcars)
data(mpg)
```
How many times does the letter “e” occur in mtcars rownames?
```{r}
sum(str_count(rownames(mtcars),"e"))
```
\qquad There are 25 times that letter "e" occurs in mtcars rownames.   
How many cars in mtcars have the brand Merc?
```{r}
sum(str_detect(rownames(mtcars),"Merc"))
```
\qquad There are 7 cars from Mercedes.   
How many cars in mpg have the brand(“manufacturer” in mpg) Merc?
```{r}
sum(str_detect(mpg$manufacturer,"merc"))
```
\qquad There are 4 cars in mpg have the brand Mercedes.  
Contrast the mileage data for Merc cars as reported in mtcars and mpg. Use tables, plots, and a short explaination.
\qquad Is it referring to mpg by mileage? 

## Problem 4
Install the babynames package.
```{r}
library(babynames)

```
Draw a sample of 500,000 rows from the babynames data
```{r}
names_samp<- babynames[sample(1:1924655, 500000),]
```
Produce a tabble that displays the five most popular boy names and girl names in the years 1880,1920, 1960, 2000.
```{r, include=TRUE}
boy_name1880<- names_samp %>% filter(sex=="M" &year==1880) %>% group_by(name, sex, year, n)
boy_name1880<- boy_name1880 %>% arrange(desc(n))
kable(boy_name1880[1:5,])

boy_name1920<- names_samp %>% filter(sex=="M" &year==1920) %>% group_by(name, sex, year, n)
boy_name1880<- boy_name1920 %>% arrange(desc(n))
kable(boy_name1920[1:5,])

boy_name1960<- names_samp %>% filter(sex=="M" &year==1960) %>% group_by(name, sex, year, n)
boy_name1960<- boy_name1960 %>% arrange(desc(n))
kable(boy_name1960[1:5,])

boy_name2000<- names_samp %>% filter(sex=="M" &year==2000) %>% group_by(name, sex, year, n)
boy_name2000<- boy_name2000 %>% arrange(desc(n))
kable(boy_name2000[1:5,])

girl_name1880<- names_samp %>% filter(sex=="F" &year==1880) %>% group_by(name, sex, year, n)
girl_name1880<- girl_name1880 %>% arrange(desc(n))
kable(girl_name1880[1:5,])

girl_name1920<- names_samp %>% filter(sex=="F" &year==1920) %>% group_by(name, sex, year, n)
girl_name1920<- girl_name1920 %>% arrange(desc(n))
kable(girl_name1920[1:5,])

girl_name1960<- names_samp %>% filter(sex=="F" &year==1960) %>% group_by(name, sex, year, n)
girl_name1960<- girl_name1960 %>% arrange(desc(n))
kable(girl_name1960[1:5,])

girl_name2000<- names_samp %>% filter(sex=="F" &year==2000) %>% group_by(name, sex, year, n)
girl_name2000<- girl_name2000 %>% arrange(desc(n))
kable(girl_name2000[1:5,])
```
What names overlap boys and girls?
```{r,include=TRUE}
overlaps_names<- names_samp %>% group_by(name,sex) %>% summarize(freq=sum(n))
overlaps_names<- overlaps_names %>% summarize(overlaps=length(sex))
overlaps_names<- overlaps_names %>% filter(overlaps==2) 
kable(overlaps_names[1:10,])
```
\qquad Here are 10 illustrations.  
What names were used in the 19th century but have not been used in the 21sth century?
\qquad 20th century? 
```{r, include=TRUE}
names_19 <- names_samp %>% filter(year<1899)
names_19<- unique(names_19$name)
names_21<- names_samp %>% filter(year>2000)
names_21<- unique(names_21$name)

diff_names<- intersect(names_19, names_21)
diff_names[1:20]
```
\qquad Here are 20 illustrations.  
Produce a chart that shows the relative frequency of the names “Donald”, “Hilary”, “Hillary”, “Joe”, “Barrack”, over the years 1880 through 2017.
```{r, include=TRUE, fig.align="center"}
spec_names<- names_samp %>% filter(year>=1880|year<=2017)
spec_names<- spec_names %>% filter(name=="Donald"|name=="Hilary"|name=="Joe"|name=="Barrack") %>% group_by(name) %>% summarize(freq=sum(n))

ggplot(spec_names)+ 
  aes(x = name, weight = freq) +
  geom_bar() +
  ylab("Frequency of having same names with President Candidates")
```



